---
output: github_document
bibliography: vicar.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# vicar: Various Ideas for Confounder Adjustment in Regression
[![Windows Build](https://ci.appveyor.com/api/projects/status/github/dcgerard/vicar?branch=master&svg=true)](https://ci.appveyor.com/project/dcgerard/vicar)
[![Linux Build](https://travis-ci.org/dcgerard/vicar.svg?branch=master)](https://travis-ci.org/dcgerard/vicar)
[![Coverage Status](https://coveralls.io/repos/github/dcgerard/vicar/badge.svg?branch=master)](https://coveralls.io/github/dcgerard/vicar?branch=master)
<!--[![Coverage](https://img.shields.io/codecov/c/github/dcgerard/vicar/master.svg)](https://codecov.io/github/dcgerard/vicar?branch=master)
[![License: GPL v3](https://img.shields.io/badge/License-GPL%20v3-blue.svg)](http://www.gnu.org/licenses/gpl-3.0) -->

R package implementing many methods for accounting for unobserved
confounding in linear regression, with or without control genes.

*Give computing environment used to test package.*

## Quick Start

*Add quick start instructions here.*

## Citing vicar

If you find that this software is useful for your research, please
cite our paper. See [here](inst/CITATION) more information.

*It is better to add this info to [inst/CITATION](inst/CITATION) and
then the citation info can be obtained by viewing this file or by
running `citation("vicar")` in R. I've added placeholder citation info
(please fix as needed). I realize that your citation info is a little
bit more complicated because you have several papers that could be
cited, but I think that can be incorporated into the CITATION file.*

## Description

Let

Y = XB + ZA + E,

for

* Y an n by p matrix of gene expression data with n samples and p genes,
* X an n by q matrix of q covariates,
* B a q by p matrix of unobserved coefficients for the observed covariates,
* Z an n by k matrix of hidden confounders,
* A a k by p matrix of hidden coefficients for the hidden confounders, and
* E an n by p matrix of independent normal errors with column variances s1,...,sp.

Not accounting for the hidden covariates, Z, can reduce power and
result in poor control of false discovery rate. This package provides
a suite of functions to adjust for hidden confounders, both when one
has and does not have access to control genes.

The functions `mouthwash` and `backwash` can adjust for hidden confounding when one
does not have access to control genes. They do so via non-parametric
empirical Bayes methods that use the powerful methodology of Adaptive
SHrinkage [@stephens2016false] within the factor-augmented regression
framework described in @wang2015confounder. `backwash` is a slightly more 
Bayesian version of `mouthwash`.

When one has control genes, there are many approaches to take. Such
methods include RUV2 [@gagnon2012using], RUV4 [@gagnon2013removing],
and CATE [@wang2015confounder]. This package adds to the field of
confounder adjustment with control genes by

1. Implementing a version of CATE that is calibrated using control genes similarly to the method in @gagnon2013removing. The function is called `vruv4`.
2. Introduces RUV3, a version of RUV that can be considered both RUV2 and RUV4. The function is called `ruv3`.
3. Introduces RUV-impute, a more general framework for accounting for hidden confounders in regression. The function is called `ruvimpute`
4. Introduces RUV-Bayes, a Bayesian version of RUV. The function is called `ruvb`.

See also the related R packages
[`cate`](https://cran.r-project.org/web/packages/cate/index.html)
[@wang2015cate] and
[`ruv`](https://cran.r-project.org/web/packages/ruv/index.html)
[@bartsch2015ruv].

Check out [NEWS.md](NEWS.md) to see what's new with each version.

## Installation
To install, run in R:

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("dcgerard/vicar")
```

## Vignettes
I've provided three vignettes to help you get started with vicar. By default, the vignettes are not built when you use `install_github`. To build the vignettes during installation, run
```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("dcgerard/vicar", build_vignettes = TRUE)
```
Note that this will result in a somewhat slower install. The first vignette, _sample\_analysis_, gives a sample analysis using vicar to account for hidden confounding. The second vignette, _customFA_, gives a few instructions on how to incorporate user-defined factor analyses with the confounder adjustment procedures implemented in vicar. The third vignette, _custom\_prior_, gives instructions and examples on incorporating a user-specified prior into `ruvb`. To see these vignettes after install, run
```{r, eval = FALSE}
utils::vignette("sample_analysis", package = "vicar")
utils::vignette("customFA", package = "vicar")
utils::vignette("custom_prior", package = "vicar")
```

## References
